---
title: "AllBC_outcomes"
author: "Ji-Qing Chen"
date: "9/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(survival)
library(survminer)
library(stargazer)
```

# load data
```{r}
load("/Users/chenjiqing/Public/bladder cancer/Basic Analysis/data_clean2.RData")
rm(Recu,phenotype1)

# load extended immune
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Bladder_Extended_cell_proportion.RData")
```

# clean data
```{r}
phenotype2 <- phenotype2 %>% mutate(RecurYN2 = ifelse(RecurYN=='yes','yes','no'),
                                    ProgressYN2 = ifelse(ProgressYN== 'yes','yes','no'))
phenotype2$RecurYN2[which(is.na(phenotype2$RecurYN2) == T)] <- ''
phenotype2$ProgressYN2[which(is.na(phenotype2$ProgressYN2) == T)] <- ''
phenotype2 <- phenotype2 %>% 
              mutate(Outcome = ifelse(RecurYN2 =="yes"|
                                      ProgressYN2 =="yes"|
                                      ddstat=="dead","yes","no")) %>%
              mutate(Outcome = ifelse(Outcome=="yes",1,0),
                     Death = ifelse(ddstat == "dead",1,0)) %>%
              mutate(OutcomeTime = ifelse(RecurYN2 =="yes", 
                                          recsurgtmd,survreftmd)) %>%
              mutate(OutcomeTime = ifelse(RecurYN2 =="no"&
                                           ProgressYN2 =="yes",
                                           progearlydxtmd,OutcomeTime))
# weird data
weird = phenotype2[which(phenotype2$ddstat=="dead" & phenotype2$RecurYN2=="yes" &
                         phenotype2$survreftmd<phenotype2$recsurgtmd),]
weird2 = weird %>% select(survreftmd,survearlydxtmd,blood_collect_age_early,blood_collect_age_ref)

# remove weird data
phenotype2_2 <- phenotype2 %>% filter(Sample_Name != weird$Sample_Name)

phenotype3 <- phenotype2_2 %>%
          mutate(SurgToBlood2 = SurgToBlood/365) %>%
          mutate(SurgToBlood3 = ifelse(SurgToBlood2<0.5,"<6month",
          ifelse(SurgToBlood2<1,"6month<=x<1year",
          ifelse(SurgToBlood2<2,"1year<=x<2year",
          ifelse(SurgToBlood2<3,"2year<=x<3year",">3year")))))%>%
          dplyr::select(Sample_Name,refage,sex,grade2,smkstat2,Death,survreftmd,
                 ImmunoTx,Outcome,OutcomeTime,SurgToBlood,SurgToBlood3,muscinv) %>%
          left_join(Cell_Type_Est) %>%
          mutate(Bas2 = round(Bas*100,1),Bmem2 = round(Bmem*100,1),
                 Bnv2 = round(Bnv*100,1),CD4mem2 = round(CD4mem*100,1),
                 CD4nv2 = round(CD4nv*100,1),CD8mem2 = round(CD8mem*100,1),
                 CD8nv2 = round(CD8nv*100,1),Eos2 = round(Eos*100,1),
                 Mono2 = round(Mono*100,1),Neu2 = round(Neu*100,1),
                 NK2 = round(NK*100,1),Treg2 = round(Treg*100,1)) %>%
          mutate(CD4 = CD4mem2 + CD4nv2,
                 CD8 = CD8mem2 + CD8nv2,
                 Bcell = Bmem2 + Bnv2) %>%
          mutate(T_cell = CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Treg2,
                 Lymph = Bmem2 + Bnv2 + NK2 + T_cell,
                 CD4T =  CD4mem2 + CD4nv2 + Treg2,
                 CD8T = CD8mem2 + CD8nv2) %>%
          mutate(Lymph = ifelse(Lymph == 0, Lymph+0.01, Lymph),
                 CD8T = ifelse(CD8T == 0, CD8T+0.01, CD8T),
                 CD4nv2 = ifelse(CD4nv2 == 0, CD4nv2+0.01, CD4nv2),
                 CD8nv2 = ifelse(CD8nv2 == 0, CD8nv2+0.01,CD8nv2)) %>%
          mutate(NLR = ifelse(Neu2/Lymph == Inf|Neu2/Lymph == -Inf,
                              NA, Neu2/Lymph),
                 MLR = ifelse(Mono2/Lymph == Inf|Mono2/Lymph == -Inf,
                              NA, Mono2/Lymph),
                 CD4_to_8 = CD4T/CD8T,
                 CD4_to_8 = ifelse(CD4_to_8 == Inf|CD4_to_8 == -Inf,
                                   NA, CD4_to_8),
                 CD4mem_to_nv = ifelse(CD4mem2/CD4nv2 == Inf|CD4mem2/CD4nv2 == -Inf, 
                                       NA, CD4mem2/CD4nv2),
                 CD8mem_to_nv = ifelse(CD8mem2/CD8nv2 == Inf|CD8mem2/CD8nv2 == -Inf, 
                                       NA, CD8mem2/CD8nv2),
                 Tmem_to_nv = ifelse((CD8mem2+CD4mem2)/(CD8nv2+CD4nv2) == Inf|
                                       (CD8mem2+CD4mem2)/(CD8nv2+CD4nv2) == -Inf, 
                                     NA, (CD8mem2+CD4mem2)/(CD8nv2+CD4nv2)))

phenotype3$SurgToBlood3 <- factor(phenotype3$SurgToBlood3, levels = c("<6month","6month<=x<1year","1year<=x<2year","2year<=x<3year",">3year"))
phenotype3$SurgToBlood3 = relevel(phenotype3$SurgToBlood3, ref = "<6month")

# Winsorization
phenotype3 <- phenotype3 %>%
         mutate(Bas2 = ifelse(Bas2 <= quantile(phenotype3$Bas2, probs = 0.98),
                              Bas2,quantile(phenotype3$Bas2, probs = 0.98)),
                Bmem2 = ifelse(Bmem2 <= quantile(phenotype3$Bmem2, probs = 0.98),
                               Bmem2,quantile(phenotype3$Bmem2, probs = 0.98)),
                Bnv2 = ifelse(Bnv2 <= quantile(phenotype3$Bnv2, probs = 0.98),
                              Bnv2,quantile(phenotype3$Bnv2, probs = 0.98)),
                CD4mem2 = ifelse(CD4mem2 <= quantile(phenotype3$CD4mem2, probs = 0.98),
                                 CD4mem2,quantile(phenotype3$CD4mem2, probs = 0.98)),
                CD4nv2 = ifelse(CD4nv2 <= quantile(phenotype3$CD4nv2, probs = 0.98),
                                CD4nv2,quantile(phenotype3$CD4nv2, probs = 0.98)),
                CD8mem2 = ifelse(CD8mem2 <= quantile(phenotype3$CD8mem2, probs = 0.98),
                                 CD8mem2,quantile(phenotype3$CD8mem2, probs = 0.98)),
                CD8nv2 = ifelse(CD8nv2 <= quantile(phenotype3$CD8nv2, probs = 0.98),
                                CD8nv2,quantile(phenotype3$CD8nv2, probs = 0.98)),
                Eos2 = ifelse(Eos2 <= quantile(phenotype3$Eos2, probs = 0.98),
                              Eos2,quantile(phenotype3$Eos2, probs = 0.98)),
                Mono2 = ifelse(Mono2 <= quantile(phenotype3$Mono2, probs = 0.98),
                               Mono2,quantile(phenotype3$Mono2, probs = 0.98)),
                Neu2 = ifelse(Neu2 <= quantile(phenotype3$Neu2, probs = 0.02),
                              quantile(phenotype3$Neu2, probs = 0.02),Neu2),
                NK2 = ifelse(NK2 <= quantile(phenotype3$NK2, probs = 0.98),
                             NK2,quantile(phenotype3$NK2, probs = 0.98)),
                Treg2 = ifelse(Treg2 <= quantile(phenotype3$Treg2, probs = 0.98),
                               Treg2,quantile(phenotype3$Treg2, probs = 0.98)))

NLR_win = quantile(phenotype3$NLR[!is.na(phenotype3$NLR)], probs = 0.98)
MLR_win = quantile(phenotype3$MLR[!is.na(phenotype3$MLR)], probs = 0.98)
CD4_to_8_win = quantile(phenotype3$CD4_to_8[!is.na(phenotype3$CD4_to_8)], 
                        probs = 0.98)
CD4mem_to_nv_win = quantile(phenotype3$CD4mem_to_nv[!is.na(phenotype3$CD4mem_to_nv)], 
                            probs = 0.98)
CD8mem_to_nv_win = quantile(phenotype3$CD8mem_to_nv[!is.na(phenotype3$CD8mem_to_nv)], 
                            probs = 0.98)
Tmem_to_nv_win = quantile(phenotype3$Tmem_to_nv[!is.na(phenotype3$Tmem_to_nv)], 
                          probs = 0.98)

phenotype4 <- phenotype3 %>%
          mutate(NLR = ifelse(NLR <= NLR_win,NLR,NLR_win),
                 MLR = ifelse(MLR <= MLR_win,MLR,MLR_win),
                 CD4_to_8 = ifelse(CD4_to_8 <= CD4_to_8_win,CD4_to_8,CD4_to_8_win),
                 CD4mem_to_nv = ifelse(CD4mem_to_nv <= CD4mem_to_nv_win,
                                       CD4mem_to_nv,CD4mem_to_nv_win),
                 CD8mem_to_nv = ifelse(CD8mem_to_nv <= CD8mem_to_nv_win,
                                       CD8mem_to_nv,CD8mem_to_nv_win),
                 Tmem_to_nv = ifelse(Tmem_to_nv <= Tmem_to_nv_win,
                                     Tmem_to_nv,Tmem_to_nv_win)) %>% 
         mutate(NLR_cut = cut(NLR, breaks = c(quantile(NLR,probs=seq(0,1, by = 0.25)))))
phenotype4$NLR_cut[which(is.na(phenotype4$NLR_cut) == T)] <- '(0.225,1.62]'
```

# check collinearity
```{r}
library(rms)
# Recurrence + progression +Death
os1 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2  + Neu2 + NK2 + Treg2, data=phenotype4)
vif(os1) # Neu has vif larger than 10
cox.zph(os1) # age, CD8nv2 violate proportional assumption
os2 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2 + NK2 + Treg2, data=phenotype4)
vif(os2) # No one has vif larger than 10
cox.zph(os2) # age and CD8nv2 violate proportional assumption, except Global

os3 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + NLR, data=phenotype4)
vif(os3) # No one has vif larger than 10
cox.zph(os3) # age and NLR violate proportional assumption
summary(os3)
os3_2 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + NLR_cut, data=phenotype4)
vif(os3_2) # No one has vif larger than 10
cox.zph(os3_2) # age violates proportional assumption

# OS
os4 <- coxph(Surv(survreftmd, Death) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2  + Neu2 + NK2 + Treg2, data=phenotype4)
vif(os4) # Neu has vif larger than 10
cox.zph(os4) # sex, muscinv, grade, Bnv, CD4mem, Neu violate proportional assumption
os5 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2 + NK2 + Treg2, data=phenotype4)
vif(os5) 
cox.zph(os5) # Bnv, CD4mem violate proportional assumption

os6 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + NLR, data=phenotype4)
vif(os6) # No one has vif larger than 10
cox.zph(os6) # No one violate proportional assumption

table(phenotype2$muscinv, phenotype2$grade)
table(phenotype2$muscinv, phenotype2$stage)
```

# Extended immune cells Recurrence + progression +Death
```{r}
os1 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2 + NK2 + Treg2, data=phenotype4)
os2 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bas2, data=phenotype4)
os3 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bmem2, data=phenotype4)
os4 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bnv2, data=phenotype4)
os5 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + CD4mem2, data=phenotype4)
os6 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + CD4nv2, data=phenotype4)
os7 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + CD8mem2, data=phenotype4)
os8 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + CD8nv2, data=phenotype4)
os9 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Eos2, data=phenotype4)
os10 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Mono2, data=phenotype4)
os11 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + NK2, data=phenotype4)
os12 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Treg2, data=phenotype4)
os13 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Neu2, data=phenotype4)
cox.zph(os8)
summary(os1)

a = list(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10,os11,os12,os13)
b = list()
for(k in 1:length(a)) {
  cc = exp(confint(a[[k]]))
  b[[k]] = cc
  print(b)
}
b

summary(a[[1]])$concordance[[1]]
c = list("c-index")
for(k in 1:length(a)) {
  ll = round(summary(a[[k]])$concordance[[1]],digits = 3)
  c[[1]][k+1] = ll
  print(c)
}
c

cox.zph(a[[8]])$table["GLOBAL",3]
d = list("p.value (proportional assumption)")
for(k in 1:length(a)) {
  lll = round(cox.zph(a[[k]])$table["GLOBAL",3],digits = 4)
  d[[1]][k+1] = lll
  print(d)
}
d

stargazer(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10,os11,os12,os13,apply.coef = exp,ci = T, ci.custom = b,p.auto=F,type = "html", out = "AllOutcome_Extcell_AllBC.htm", star.cutoffs=c(0.05,0.01,0.001), column.labels = c("Multivariable model"),column.separate = c(13), 
          covariate.labels = c("Age","Female (ref.= Male)","Muscle Invasive","Grade 3 + 4 (ref.= Grade 1 + 2)","Ever-smoker (ref.= Non-smoker)","BCG Trt. (ref.= Without BCG Trt.)" ,"Basophils","Bmem","Bnv","CD4mem","CD4nv","CD8mem","CD8nv","Eosinophil","Monocyte","NK","Treg","Neutrophil"), 
          dep.var.labels.include = FALSE, add.lines=c(c,d), omit.stat = c("max.rsq","ll","wald","logrank"),keep.stat = c("n","rsq","aic"),dep.var.caption = "Cox proportional hazards recurrence-free survival models",
          notes = c("Outcommes: recurrence, progression, and death. Winsorization was used for 2% value of each immune cell profile. Age and CD8nv violates proportional assumption in all model."),notes.label = "",notes.align = "l")
```

# Extended immune cells OS
```{r}
os1 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2 + NK2 + Treg2, data=phenotype4)
# Bnv, CD4mem violate proportional assumption
os2 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Bas2, data=phenotype4)
os3 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Bmem2, data=phenotype4)
os4 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Bnv2, data=phenotype4)
# Bnv violates proportional assumption
os5 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + CD4mem2, data=phenotype4)
# CD4mem violates proportional assumption
os6 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + CD4nv2, data=phenotype4)
os7 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + CD8mem2, data=phenotype4)
# CD8mem violates proportional assumption
os8 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + CD8nv2, data=phenotype4)
os9 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Eos2, data=phenotype4)
os10 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Mono2, data=phenotype4)
os11 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + NK2, data=phenotype4)
# NK violates proportional assumption
os12 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Treg2, data=phenotype4)
os13 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Neu2, data=phenotype4)
# Neu violates proportional assumption
cox.zph(os13)
summary(os1)

a = list(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10,os11,os12,os13)
b = list()
for(k in 1:length(a)) {
  cc = exp(confint(a[[k]]))
  b[[k]] = cc
  print(b)
}
b

summary(a[[1]])$concordance[[1]]
c = list("c-index")
for(k in 1:length(a)) {
  ll = round(summary(a[[k]])$concordance[[1]],digits = 3)
  c[[1]][k+1] = ll
  print(c)
}
c

cox.zph(a[[8]])$table["GLOBAL",3]
d = list("p.value (proportional assumption)")
for(k in 1:length(a)) {
  lll = round(cox.zph(a[[k]])$table["GLOBAL",3],digits = 4)
  d[[1]][k+1] = lll
  print(d)
}
d

stargazer(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10,os11,os12,os13,apply.coef = exp,ci = T, ci.custom = b,p.auto=F,type = "html", out = "OSover10year_Extcell_AllBC.htm", star.cutoffs=c(0.05,0.01,0.001), column.labels = c("Multivariable model"),column.separate = c(13), 
          covariate.labels = c("Age", "Grade 3 + 4 (ref.= Grade 1 + 2)","Ever-smoker (ref.= Non-smoker)", "BCG Trt. (ref.= Without BCG Trt.)", "Basophils","Bmem","Bnv","CD4mem","CD4nv","CD8mem","CD8nv","Eosinophil","Monocyte","NK","Treg","Neutrophil"), 
          dep.var.labels.include = FALSE, add.lines=c(c,d), omit.stat = c("max.rsq","ll","wald","logrank"),keep.stat = c("n","rsq","aic"),dep.var.caption = "Cox proportional hazards overall survival models",
          notes = c("Winsorization was used for 2% value of each immune cell profile. Bnv, CD4mem, CD8mem, NK and Neu violate proportional assumption"),notes.label = "",notes.align = "l")
```

# correlation matrix: SurgToBlood, Mono, Bmem, Bnv, CD4mem, CD4nv, CD8mem, CD8nv, NK, Neu (with winsorization in immune cell proportion)
```{r}
library(PerformanceAnalytics)
A <- NMIBC3 %>%
     select(SurgToBlood,Bmem2,Bnv2,CD4mem2,CD4nv2,CD8mem2,CD8nv2,Mono2,NK2,Neu2)
chart.Correlation(A, histogram = TRUE, pch = 19)
```

# NLR, NLR cut, All CD4, All CD8, All Bcell
```{r}
os1 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + NLR, data=phenotype4)
os2 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + CD4, data=phenotype4)
# CD4 violates proportional assumption
os3 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + CD8, data=phenotype4)
# CD8 violates proportional assumption
os4 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + Bcell, data=phenotype4)
# Bcell violates proportional assumption
os5 <- coxph(Surv(survreftmd, Death) ~ refage + strata(sex) + strata(muscinv) + grade2 + smkstat2 + ImmunoTx + NLR_cut, data=phenotype4)

os6 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + NLR, data=phenotype4)
# Age and NLR violate proportional assumption
os7 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + CD4, data=phenotype4) # Age
os8 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + CD8, data=phenotype4) # Age
os9 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + Bcell, data=phenotype4) # Age
os10 <- coxph(Surv(OutcomeTime, Outcome) ~ refage + sex + muscinv + grade2 + smkstat2 + ImmunoTx + NLR_cut, data=phenotype4) # Age
cox.zph(os10)
summary(os1)

a = list(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10)
b = list()
for(k in 1:length(a)) {
  cc = exp(confint(a[[k]]))
  b[[k]] = cc
  print(b)
}

c = list("c-index")
for(k in 1:length(a)) {
  ll = round(summary(a[[k]])$concordance[[1]],digits = 3)
  c[[1]][k+1] = ll
  print(c)
}
c

d = list("p.value (proportional assumption)")
for(k in 1:length(a)) {
  lll = round(cox.zph(a[[k]])$table["GLOBAL",3],digits = 4)
  d[[1]][k+1] = lll
  print(d)
}
d

stargazer(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10,apply.coef = exp,ci = T, ci.custom = b,p.auto=F,type = "html", out = "OutcomesNot10year_NLR_CD48_AllBC.htm", star.cutoffs=c(0.05,0.01,0.001), column.labels = c("OS","RFS"),column.separate = c(5,5), 
          covariate.labels = c("Age", "Female (ref.= Male)", "Muscle Invasive","Grade 3 + 4 (ref.= Grade 1 + 2)","Ever-smoker (ref.= Non-smoker)", "BCG Trt. (ref.= Without BCG Trt.)", "NLR","CD4T","CD8T","B cell","NLR (1.62,2.38]","NLR (2.38,3.86]","NLR (3.86,14.2]"), 
          dep.var.labels.include = FALSE, add.lines=c(c,d), omit.stat = c("max.rsq","ll","wald","logrank"),keep.stat = c("n","rsq","aic"),dep.var.caption = "Cox proportional hazards Outcome models",
          notes = c("Winsorization was used for 2% value of NLR. CD4T, CD8T and B cell violate proportional assumption in OS. Age and NLR violate proportional assumption in RFS"),notes.label = "",notes.align = "l")
```

_________________________________________________________
# BD to censoring/event time (create variable)
```{r}
NMIBC3 <- NMIBC3 %>% mutate(BDtoCenorEve_Surv = TenYearSurv*30-SurgToBlood,
                            BDtoCenorEve_RFS = TenYearRFS*30-SurgToBlood)
hist(NMIBC3$BDtoCenorEve_Surv, xlab = 'BD to Censoring or Event (Days)')
hist(NMIBC3$BDtoCenorEve_RFS, xlab = 'BD to Censoring or Event (Days)')
```

# check collinearity
```{r}
library(rms)
# RFS
os1 <- coxph(Surv(TenYearRFS, TenRFS) ~ refage + strata(sex) + grade2 + smkstat2 + strata(ImmunoTx) + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2  + NK2 + Treg2 + BDtoCenorEve_RFS, data=NMIBC3)
vif(os1) 
cox.zph(os1)

# OS
os3 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2  + Neu2 + NK2 + Treg2 + BDtoCenorEve_Surv, data=NMIBC3)
vif(os3) # Neu has vif larger than 10
cox.zph(os3) # Bas violate proportional assumption
os4 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2 + NK2 + Treg2 + BDtoCenorEve_Surv, data=NMIBC3)
vif(os4) # Neu has vif larger than 10
cox.zph(os4) # Bnv,CD4mem,CD4nv,CD8nv,and NK violate proportional assumption
summary(os4)
```

# Extended immune cells OS (BDtoCenorEve_Surv)
```{r}
os1 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Bas2 + Bmem2 + Bnv2 + CD4mem2 + CD4nv2 + CD8mem2 + CD8nv2 + Eos2 + Mono2  + NK2 + Treg2+ BDtoCenorEve_Surv, data=NMIBC3)
os2 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Bas2+ BDtoCenorEve_Surv, data=NMIBC3)
os3 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Bmem2+ BDtoCenorEve_Surv, data=NMIBC3)
os4 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Bnv2+ BDtoCenorEve_Surv, data=NMIBC3)
os5 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + CD4mem2+ BDtoCenorEve_Surv, data=NMIBC3)
os6 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + CD4nv2+ BDtoCenorEve_Surv, data=NMIBC3)
os7 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + CD8mem2+ BDtoCenorEve_Surv, data=NMIBC3)
os8 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + CD8nv2+ BDtoCenorEve_Surv, data=NMIBC3)
os9 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Eos2+ BDtoCenorEve_Surv, data=NMIBC3)
os10 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Mono2+ BDtoCenorEve_Surv, data=NMIBC3)
os11 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + NK2+ BDtoCenorEve_Surv, data=NMIBC3)
os12 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Treg2+ BDtoCenorEve_Surv, data=NMIBC3)
os13 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + Neu2+ BDtoCenorEve_Surv, data=NMIBC3)
os14 <- coxph(Surv(TenYearSurv, TenDead) ~ refage + sex + grade2 + smkstat2 + ImmunoTx + BDtoCenorEve_Surv, data=NMIBC3)
cox.zph(os1)
summary(os1)

a = list(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10,os11,os12,os13,os14)
b = list()
for(k in 1:length(a)) {
  cc = exp(confint(a[[k]]))
  b[[k]] = cc
  print(b)
}
b

summary(a[[1]])$concordance[[1]]
c = list("c-index")
for(k in 1:length(a)) {
  ll = round(summary(a[[k]])$concordance[[1]],digits = 3)
  c[[1]][k+1] = ll
  print(c)
}
c

cox.zph(a[[8]])$table["GLOBAL",3]
d = list("p.value (proportional assumption)")
for(k in 1:length(a)) {
  lll = round(cox.zph(a[[k]])$table["GLOBAL",3],digits = 4)
  d[[1]][k+1] = lll
  print(d)
}
d

stargazer(os1,os2,os3,os4,os5,os6,os7,os8,os9,os10,os11,os12,os13,os14,apply.coef = exp,ci = T, ci.custom = b,p.auto=F,type = "html", out = "OS_Extcell_BDtoCenorEve_Multi.htm", star.cutoffs=c(0.05,0.01,0.001), column.labels = c("Multivariable model"),column.separate = c(14), 
          covariate.labels = c("Age", "Female (ref.= Male)", "Grade 3 + 4 (ref.= Grade 1 + 2)","Ever-smoker (ref.= Non-smoker)", "BCG Trt. (ref.= Without BCG Trt.)", "Basophils","Bmem","Bnv","CD4mem","CD4nv","CD8mem","CD8nv","Eosinophil","Monocyte","NK","Treg","Neutrophil","BDtoCenOrEve"), 
          dep.var.labels.include = FALSE, add.lines=c(c,d), omit.stat = c("max.rsq","ll","wald","logrank"),keep.stat = c("n","rsq","aic"),dep.var.caption = "Cox proportional hazards 10-year overall survival models",
          notes = c("Winsorization was used for 2% value of each immune cell profile. Reference of blood draw time(BDtime): <6month. Bnv,CD4mem,CD4nv,CD8nv,and NK violate proportional assumption"),notes.label = "",notes.align = "l")
```