---
title: "Age_acc_partDSA"
author: "Ji-Qing Chen"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MST)
library(dplyr)
library(survival)
library(survminer)
library(partDSA)

library(ggsci)
library(cowplot)
```

# load data
```{r}
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/BLCR_classed_cellprop_5FoldCross.RData") # immune cell already winsorized
```

# imm + clinical features (OS)
```{r}
#first subset dataset with those variables you want to include in the tree
xinfer_imm<-subset(BLCR_classed_cellprop, select =c(Age, hannum_acc, Sex, Tumor_grade, Smoking_status, BCG, Bmem,CD4mem,CD8mem,CD8nv,Neu,Treg,NK,Bas))
OSTime = BLCR_classed_cellprop$TenYearSurv
OScen = BLCR_classed_cellprop$TenDead
model.IPCW_imm = partDSA(x=xinfer_imm, y=Surv(log(OSTime),OScen),
                     control = DSA.control(loss.function = "IPCW"))
model.IPCW_imm
showDSA(model.IPCW_imm)
#Best 3 partitions
 # Partition 1 [of 3]:
  #  (CD8nv <= 1.760000) && (Neu <= 76.460000)
  #Partition 2 [of 3]:
   # (76.460000 < Neu)
  #Partition 3 [of 3]:
   # (1.760000 < CD8nv) && (Neu <= 76.460000)

GroupOS <- BLCR_classed_cellprop %>% 
           mutate(Group = ifelse(Neu > 76.46, "Group 2", 
                          ifelse((Neu <= 76.46) & (CD8nv <= 1.76),"Group 1","Group 3")))
table(GroupOS$Group)
table(GroupOS$Class, GroupOS$Group)

# KM plot
fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  legend = "top" ,legend.title = "", 
  legend.labs = c("Group 1","Group 2","Group 3"),
  palette = "nejm",
  conf.int = T,
  surv.median.line = "hv"
  )

surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value
GroupOS2 <- GroupOS %>% filter(Group == "Group 1" | Group == "Group 2")
surv_diff2 <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS2)
pchisq(surv_diff2$chisq, length(surv_diff2$n)-1, lower.tail = FALSE)
GroupOS3 <- GroupOS %>% filter(Group == "Group 1" | Group == "Group 3")
surv_diff3 <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS3)
pchisq(surv_diff3$chisq, length(surv_diff3$n)-1, lower.tail = FALSE)

summary(os <- coxph(Surv(TenYearSurv, TenDead) ~ Group, data=GroupOS))
OS_10y$plot <- OS_10y$plot + 
          annotate("text", x = 66, y = 0, label = 61.67) +
          annotate("text", x = 10, y = 0.3, label = "Group 2 \n (referent group: Group 1) \n HR = 4.93 \n C.I. = 2.78 - 8.71 \n P = 6.27E-10",size = 3) +
          annotate("text", x = 40, y = 0.3, label = "Group 3 \n (referent group: Group 1) \n HR = 0.54 \n C.I. = 0.35 - 0.85 \n P = 0.006",size = 3) +
          theme(plot.title = element_text(size = 10))

# Cox model
Cox_G1 = coxph(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
Cox_G2 = coxph(Surv(TenYearSurv, TenDead) ~ Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Class + Group, data = GroupOS)
Cox_G3 = coxph(Surv(TenYearSurv, TenDead) ~ Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Class + Group + Bnv + CD4mem + CD8mem + NK + CD4nv + Neu, data = GroupOS)

library(rms)
vif(Cox_G3)
summary(Cox_G1)
summary(Cox_G2)
cox.zph(Cox_G1)
```

# partDSA vs SS-RPMM class membership
```{r}
# A
table(GroupOS$Class, GroupOS$Group)
GroupOS2 <- GroupOS %>% mutate(group_class = ifelse(Group=='Group 1' & Class =='rL','G1_rL',
                                                    ifelse(Group=='Group 1' & Class =='rR','G1_rR',
                                                           ifelse(Group=='Group 2','G2',
                                                                  ifelse(Group=='Group 3' & Class =='rL','G3_rL','G3_rR')))))
table(GroupOS2$group_class)

fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ group_class, data = GroupOS2)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  #ylim = c(0.4,1),
  legend = "top" ,legend.title = "", 
  legend.labs = c('G1_rL', 'G1_rR', 'G2', 'G3_rL', 'G3_rR'),
  linetype =c(4,1,1,4,1),
  palette = c("#bc3c29","#bc3c29","#0072b5","#e18727","#e18727"),
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_10y)

OS_10y$plot <- OS_10y$plot + 
          annotate("text", x = 66.2, y = 0, label = 61.67) +
          annotate("text", x = 10, y = 0.3, label = "G1_rR \n (referent group: G1_rL) \n HR = 0.36 \n C.I. = 0.25 - 0.52 \n P = 1.54E-8", size = 3) +
          annotate("text", x = 37, y = 0.3, label = "G3_rR \n (referent group: G3_rL) \n HR = 0.33 \n C.I. = 0.15 - 0.76 \n P = 0.006", size = 3) +
          theme(plot.title = element_text(size = 10))

GroupOS3 <- GroupOS2 %>% filter(group_class == "G1_rL" | group_class == "G1_rR")
summary(os <- coxph(Surv(TenYearSurv, TenDead) ~ group_class, data=GroupOS3))
surv_diff2 <- survdiff(Surv(TenYearSurv, TenDead) ~ group_class, data = GroupOS3)
pchisq(surv_diff2$chisq, length(surv_diff2$n)-1, lower.tail = FALSE)

GroupOS4 <- GroupOS2 %>% filter(group_class == "G3_rL" | group_class == "G3_rR")
summary(os <- coxph(Surv(TenYearSurv, TenDead) ~ group_class, data=GroupOS4))
surv_diff2 <- survdiff(Surv(TenYearSurv, TenDead) ~ group_class, data = GroupOS4)
pchisq(surv_diff2$chisq, length(surv_diff2$n)-1, lower.tail = FALSE)

# B
table(GroupOS$Class, GroupOS$Group)
GroupOS2 <- GroupOS %>% mutate(group_class = ifelse(Group=='Group 1' & Class =='rL','G1_rL',
                                                    ifelse(Group=='Group 1' & Class =='rR','G1_rR',
                                                           ifelse(Group=='Group 2','G2','G3'))))
table(GroupOS2$group_class)

fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ group_class, data = GroupOS2)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  legend = "top" ,legend.title = "", 
  legend.labs = c('G1_rL', 'G1_rR', 'G2', 'G3'),
  linetype =c(4,1,1,1),
  palette = c("#bc3c29","#bc3c29","#0072b5","#e18727"),
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_10y)

summary(os <- coxph(Surv(TenYearSurv, TenDead) ~ group_class, data=GroupOS2))
GroupOS3 <- GroupOS2 %>% filter(group_class == "G1_rL" | group_class == "G2")
surv_diff2 <- survdiff(Surv(TenYearSurv, TenDead) ~ group_class, data = GroupOS3)
pchisq(surv_diff2$chisq, length(surv_diff2$n)-1, lower.tail = FALSE)

GroupOS4 <- GroupOS2 %>% filter(group_class == "G1_rL" | group_class == "G3")
surv_diff2 <- survdiff(Surv(TenYearSurv, TenDead) ~ group_class, data = GroupOS4)
pchisq(surv_diff2$chisq, length(surv_diff2$n)-1, lower.tail = FALSE)

OS_10y$plot <- OS_10y$plot + 
          annotate("text", x = 110, y = 0, label = 115.98) +
          annotate("text", x = 20, y = 0.3, label = "G2 (referent group: G1_rL) \n HR = 4.38 \n C.I. = 2.38 - 8.07 \n P = 3.67E-7") +
          annotate("text", x = 70, y = 0.3, label = "G3 (referent group: G1_rL) \n HR = 0.97 \n C.I. = 0.51 - 1.84 \n P = 0.92") +
          theme(plot.title = element_text(size = 10))
```

# 0 - 60 month for group 1 and 3
```{r}
library(tidyverse)
load("/Users/chenjiqing/Public/bladder cancer/Basic Analysis/data_clean2.RData")
rm(Recu,phenotype1)
phenotype2 <- phenotype2 %>%
              select(Sample_Name, fiveDead, fiveYearSurv) %>%
              `colnames<-`(c('Sample_ID','fiveDead','fiveYearSurv'))

GroupOS2 <- GroupOS %>%
            filter(Group == "Group 1" | Group == "Group 3") %>%
            rownames_to_column(var = 'Sample_ID') %>%
            left_join(phenotype2)

fit_5y = survfit(Surv(fiveYearSurv, fiveDead) ~ Group, data = GroupOS2)
OS_5y <- ggsurvplot(
  fit_5y,
  xlab = "Months", 
  ylab = "5-year overall survival probability",
  ylim = c(0.80,1),
  legend = "top" ,legend.title = "", 
  legend.labs = c("Group 1","Group 3"),
  palette = c("#bc3c29","#e18727"),
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_5y)
surv_diff <- survdiff(Surv(fiveYearSurv, fiveDead) ~ Group, data = GroupOS2)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE)
summary(os <- coxph(Surv(fiveYearSurv, fiveDead) ~ Group, data=GroupOS2))

OS_5y$plot <- OS_5y$plot + 
          annotate("text", x = 20, y = 0.9, label = "Group 3 (referent group: Group 1) \n HR = 0.85 \n C.I. = 0.47 - 1.57 \n P = 0.61") +
          theme(plot.title = element_text(size = 10))
```

# 61 month - 120 month for group 1 and 3
```{r}
GroupOS_2 <- GroupOS %>%
             filter(Group == "Group 1" | Group == "Group 3") %>%
             filter(TenYearSurv > 60)
table(GroupOS_2$Class, GroupOS_2$Group)

# KM plot
fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS_2)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  xlim = c(60,121),
  ylim = c(0.7,1),
  legend = "top" ,legend.title = "", 
  legend.labs = c("Group 1","Group 3"),
  palette = c("#bc3c29","#e18727"),
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS_2)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value
summary(os <- coxph(Surv(TenYearSurv, TenDead) ~ Group, data=GroupOS_2))

OS_10y$plot <- OS_10y$plot + 
          annotate("text", x = 75, y = 0.8, label = "Group 3 (referent group: Group 1) \n HR = 0.37 \n C.I. = 0.19 - 0.71 \n P = 0.002") +
          theme(plot.title = element_text(size = 10))

Cox_G1 = coxph(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS_2)
library(rms)
vif(Cox_G1)
summary(Cox_G1)
cox.zph(Cox_G1)
```

# immune cell type proportion difference
```{r}
# cell proportion
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Bladder_Extended_cell_proportion2.RData")

NMIBC_immune <- GroupOS %>% 
                rownames_to_column(., var = 'Sample_Name') %>% 
                select(Sample_Name, Group) %>%
                left_join(Extended2) %>%
                mutate(T_cell = CD4mem + CD4nv + CD8mem + CD8nv + Treg, # not show this one
                       Lymph = Bmem + Bnv + NK + T_cell) %>%
                mutate(NLR = ifelse(Lymph == 0,Neu/0.01, Neu/Lymph)) %>%
                select(-T_cell,-Lymph)
```

# Function
```{r}
# Plot Cell Type
Plot_Cell_Type <- function(DF, YLab, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = Cell_Type, y = Percentage, 
                                      by = Group, color = Group)) +
                       geom_boxplot(position = position_dodge(width = 0.95), 
                                    outlier.shape = NA, 
                                    outlier.colour = NA, 
                                    outlier.size = NA) +
                      geom_point(position = position_jitterdodge(dodge.width= 0.95, 
                                                                 jitter.width = 0.25), 
                                 
                                  alpha=0.5,size = 0.1) +
                      stat_compare_means(aes(group = Group), 
                                         label = "p.format", 
                                         label.x = P_X, label.y = P_Y, 
                                         size = 3.6, 
                                         method = "anova") +
                      labs(x = "", y = YLab, colour = "") +
                      theme_classic() +
                      scale_color_nejm() +
                      theme(legend.position = "bottom", # legend.position controls where is legend
                            text=element_text(size=12, 
                                              face = "bold"),
                            plot.margin = unit(c(6,0,6,0), "pt")) + 
                     scale_y_continuous(labels = scales::percent_format(accuracy = 1)) 
                return(PLOT)
}

NeuPlot <- function(DF){
            ggplot(DF, aes(x = "Neu",y = Neu, 
                           by = Group, color = Group)) +
            geom_boxplot(position = position_dodge(width = 0.95), 
                         outlier.shape = NA, 
                         outlier.colour = NA, 
                         outlier.size = NA) + 
            geom_point(position = position_jitterdodge(dodge.width= 0.95, 
                                                       jitter.width = 0.25),
                       alpha=0.5,size = 0.1) + 
            stat_compare_means(aes(group = Group), 
                               label = "p.format", 
                               label.x = 0, label.y = 1, 
                               size = 3.6, 
                               method = "anova") +
            labs(x = "", y = "Cell Type Percent") +
            theme_classic() +
            scale_color_nejm() +
            theme(legend.position = "none", # legend.position controls where is legend
                  text=element_text(size=12,face = "bold"),
                  plot.margin = unit(c(6,0,6,0), "pt")) +
            scale_y_continuous(labels = scales::percent_format(accuracy = 1))
}

# Plot Cell Type
Cell_Ratio_Plot <- function(DF, YLab, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = Group, y = NLR, 
                                      by = Group, color = Group)) +
                       geom_boxplot(position = position_dodge(width = 0.95), 
                                    outlier.shape = NA, 
                                    outlier.colour = NA, 
                                    outlier.size = NA) +
                       geom_point(position = position_jitterdodge(dodge.width= 0.95, 
                                                       jitter.width = 0.25),
                                 alpha=0.5,size = 0.1) +
                       stat_compare_means(aes(group = Group), 
                                          label = "p.format", 
                                          label.x = P_X, 
                                          label.y = P_Y, 
                                          size = 3.6, 
                                          method = "anova") +
                      labs(x = "NLR", y = YLab, colour = "") +
                      theme_classic() +
                      scale_color_nejm() +
                      theme(legend.position = "none",
                            text=element_text(size=12, face = "bold"),
                            plot.margin = unit(c(6,0,6,0), "pt"), 
                            axis.text.x = element_text(size = 7))
                return(PLOT)
                
        }
```

# plot
```{r}
Estimate <- NMIBC_immune %>%
            dplyr::select(Group, Bas, Bmem, Bnv, Eos, Mono, NK, Treg) %>%
            gather(key = Cell_Type, value = Percentage, -Group) %>%
            mutate(Percentage = Percentage/100)
plot1 = Plot_Cell_Type(Estimate, YLab = "Cell Type Percent", P_X = 0, P_Y = .21)

Estimate2 <- NMIBC_immune %>%
             dplyr::select(Group, CD4mem, CD4nv, CD8mem, CD8nv) %>%
             gather(key = Cell_Type, value = Percentage, -Group) %>%
             mutate(Percentage = Percentage/100)
plot2 = Plot_Cell_Type(Estimate2, YLab = "Cell Type Percent", P_X = 0, P_Y = .4)

PlotNeu <- NMIBC_immune %>% mutate(Neu = Neu/100) %>% NeuPlot()

Ratios <- NMIBC_immune %>%
          dplyr::select(Group, NLR)
PlotNLR <- Cell_Ratio_Plot(Ratios,YLab = 'Ratio', P_X = 2, P_Y = 33)

# save
Row1 = plot_grid(plot1, ncol = 1)
Row2 = plot_grid(PlotNeu, plot2, PlotNLR, ncol = 3, rel_widths = c(1,3,1))
Comb_Plot = plot_grid(Row1, Row2, ncol = 1, rel_heights = c(1, 1))
ggsave(filename = paste0("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/Age_acc/plot/NMIBC_summary proportion2.pdf"), plot = Comb_Plot, height = 9, width = 7.5)
```

# age and age acceleration
```{r}
NMIBC_age <- GroupOS %>% 
             select(Age, mAge_Hannum, hannum_acc, Group) %>%
             `colnames<-`(c('Age','Hannum_mAge', 'Age_Acceleration', 'Group'))
```

# Function
```{r}
# Plot Cell Type
Plot_Cell_Type <- function(DF, YLab, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = Ages, y = Values, 
                                      by = Group, color = Group)) +
                       geom_boxplot(position = position_dodge(width = 0.95), 
                                    outlier.shape = NA, 
                                    outlier.colour = NA, 
                                    outlier.size = NA) +
                      geom_point(position = position_jitterdodge(dodge.width= 0.95, 
                                                                 jitter.width = 0.25), 
                                 alpha=0.5,size = 0.1) +
                      stat_compare_means(aes(group = Group), 
                                         label = "p.format", 
                                         label.x = P_X, label.y = P_Y, 
                                         size = 3.6, 
                                         method = "anova") +
                      labs(x = "", y = YLab, colour = "") +
                      theme_classic() +
                      scale_color_nejm() +
                      theme(legend.position = "bottom", # legend.position controls where is legend
                            text=element_text(size=12, 
                                              face = "bold"),
                            plot.margin = unit(c(6,0,6,0), "pt"))
                return(PLOT)
}

```

# plot
```{r}
Estimate <- NMIBC_age %>%
            dplyr::select(Group, Age, Hannum_mAge) %>%
            gather(key = Ages, value = Values, -Group) 
plot1 = Plot_Cell_Type(Estimate, YLab = "Years", P_X = 0, P_Y = 97)

Estimate2 <- NMIBC_age %>%
            dplyr::select(Group, Age_Acceleration) %>%
            gather(key = Ages, value = Values, -Group) 
plot2 = Plot_Cell_Type(Estimate2, YLab = "Years", P_X = 0, P_Y = 24)

# save
Comb_Plot = plot_grid(plot1, plot2, ncol = 1, rel_heights = c(1, 1))
ggsave(filename = paste0("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/Age_acc/plot/NMIBC_age_and_mage2.pdf"), plot = Comb_Plot, height = 9, width = 7.5)
```

# Tmem Tnv + clinical features (OS)
```{r}
# cell proportion
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Bladder_Extended_cell_proportion.RData")
Extended3 <- Extended2 %>%
             mutate(Tmem = CD4mem + CD8mem,
                    Tnv = CD4nv + CD8nv) %>%
             select(Sample_Name, Tmem, Tnv)
Extended3 <- Extended3 %>%
             mutate(Tmem = ifelse(Tmem <= quantile(Extended3$Tmem, probs = 0.98),
                              Tmem,quantile(Extended3$Tmem, probs = 0.98)),
                    Tnv = ifelse(Tnv <= quantile(Extended3$Tnv, probs = 0.98),
                              Tnv,quantile(Extended3$Tnv, probs = 0.98)))
             
BLCR_classed_cellprop2 <- BLCR_classed_cellprop %>%
                          rownames_to_column(., var = "Sample_Name") %>%
                          left_join(Extended3) %>%
                          column_to_rownames(., var = "Sample_Name")

#first subset dataset with those variables you want to include in the tree
xinfer_imm<-subset(BLCR_classed_cellprop2, select =c(Age, hannum_acc, Sex, Tumor_grade, Smoking_status, BCG, Bnv, Tmem, Tnv, Neu, NK, Class))
OSTime = BLCR_classed_cellprop2$TenYearSurv
OScen = BLCR_classed_cellprop2$TenDead
model.IPCW_imm = partDSA(x=xinfer_imm, y=Surv(log(OSTime),OScen),
                     control = DSA.control(loss.function = "IPCW"))
model.IPCW_imm
showDSA(model.IPCW_imm)
#Best 3 partitions
 # Partition 1 [of 3]:
  #  (Tnv <= 3.630000) && (Neu <= 67.090000)
  #Partition 2 [of 3]:
   # (67.090000 < Neu)
  #Partition 3 [of 3]:
   # (3.630000 < Tnv) && (Neu <= 67.090000)

GroupOS <- BLCR_classed_cellprop2 %>% 
           mutate(Group = ifelse(Neu > 67.09, "Group 2", 
                          ifelse((Neu <= 67.09) & (Tnv <= 3.63),"Group 1","Group 3")))
table(GroupOS$Group)
table(GroupOS$Class, GroupOS$Group)

# KM plot
fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  legend = "top" ,legend.title = "", 
  legend.labs = c("Group 1","Group 2","Group 3"),
  palette = "nejm",
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value
#summary(os4 <- coxph(Surv(TenYearRFS, TenRFS) ~ age_cut, data=Age_Analysis))
OS_10y$plot <- OS_10y$plot + 
          annotate("text", x = 110, y = 0, label = 115.98) +
          theme(plot.title = element_text(size = 10))

# Cox model
Cox_G1 = coxph(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
Cox_G2 = coxph(Surv(TenYearSurv, TenDead) ~ Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Class + Group, data = GroupOS)
Cox_G3 = coxph(Surv(TenYearSurv, TenDead) ~ Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Class + Group + Bnv + CD4mem + CD8mem + NK + CD4nv + Neu, data = GroupOS)

library(rms)
vif(Cox_G1)
summary(Cox_G1)
cox.zph(Cox_G1)
```

______________________________________________________________
# Not include Neu partDSA (OS)
```{r}
#first subset dataset with those variables you want to include in the tree
xinfer_imm<-subset(BLCR_classed_cellprop, select =c(Age, hannum_acc, Sex, Tumor_grade, Smoking_status, BCG, Bnv, CD4mem, CD4nv, CD8mem, NK, Class))
OSTime = BLCR_classed_cellprop$TenYearSurv
OScen = BLCR_classed_cellprop$TenDead
model.IPCW_imm = partDSA(x=xinfer_imm, y=Surv(log(OSTime),OScen),
                     control = DSA.control(loss.function = "IPCW"))
model.IPCW_imm

# Best 4 partitions
# Partition 1 [of 4]:
 #   (CD4mem <= 4.040000) && (NK <= 3.080000)
  #Partition 2 [of 4]:
   # (4.040000 < CD4mem) && (CD8mem <= 1.470000)
#  Partition 3 [of 4]:
 #   (4.040000 < CD4mem) && (1.470000 < CD8mem)
#  Partition 4 [of 4]:
 #   (CD4mem <= 4.040000) && (3.080000 < NK)

GroupOS <- BLCR_classed_cellprop %>% 
           mutate(Group = ifelse((4.04 < CD4mem) & (CD8mem > 1.47), "Group 3", 
                          ifelse((4.04 < CD4mem) & (CD8mem <= 1.47),"Group 2",
                          ifelse((CD4mem <= 4.04) & (NK > 3.08), "Group 4","Group 1"))))
table(GroupOS$Group)
table(GroupOS$Class, GroupOS$Group)

# KM plot
fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  legend = "top" ,legend.title = "", 
  legend.labs = c("Group 1","Group 2","Group 3","Group 4"),
  palette = "nejm",
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value

# Cox model
Cox_G1 = coxph(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
Cox_G2 = coxph(Surv(TenYearSurv, TenDead) ~ Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Class + Group, data = GroupOS)
Cox_G3 = coxph(Surv(TenYearSurv, TenDead) ~ Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Class + Group + Bnv + CD4nv, data = GroupOS)
summary(Cox_G3)

# Merge group 4 and group 3
GroupOS2 <- GroupOS %>%
            mutate(Group = ifelse(Group == "Group 4","Group 3",Group))
table(GroupOS2$Group)
table(GroupOS2$Class, GroupOS2$Group)

fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS2)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  legend = "top" ,legend.title = "", 
  legend.labs = c("Group 1","Group 2","Group 3"),
  palette = "nejm",
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS2)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value

# Cox model
Cox_G1 = coxph(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS2)
summary(Cox_G1)
```

______________________________________________________________
# Not include immune partDSA (OS)
```{r}
#first subset dataset with those variables you want to include in the tree
xinfer_imm<-subset(BLCR_classed_cellprop, select =c(Age, hannum_acc, Sex, Tumor_grade, Smoking_status, BCG, Class))
OSTime = BLCR_classed_cellprop$TenYearSurv
OScen = BLCR_classed_cellprop$TenDead
model.IPCW_imm = partDSA(x=xinfer_imm, y=Surv(log(OSTime),OScen),
                     control = DSA.control(loss.function = "IPCW"))
model.IPCW_imm

#Best 2 partitions
 # Partition 1 [of 2]:
  #  (hannum_acc <= -0.585271)
  #Partition 2 [of 2]:
   # (-0.585271 < hannum_acc)

GroupOS <- BLCR_classed_cellprop %>% 
           mutate(Group = ifelse(hannum_acc <= -0.585271, "Group 1", "Group 2"))
table(GroupOS$Group)
table(GroupOS$Class, GroupOS$Group)

# KM plot
fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival probability",
  legend = "top" ,legend.title = "", 
  legend.labs = c("Group 1","Group 2"),
  ylim = c(0.5,1),
  palette = "nejm",
  conf.int = T,
  surv.median.line = "hv"
  )
surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value
#summary(os4 <- coxph(Surv(TenYearRFS, TenRFS) ~ age_cut, data=Age_Analysis))

# Cox model
Cox_G1 = coxph(Surv(TenYearSurv, TenDead) ~ Group, data = GroupOS)
Cox_G2 = coxph(Surv(TenYearSurv, TenDead) ~ Age + Sex + Tumor_grade + Smoking_status + BCG + Class + Group, data = GroupOS)
summary(Cox_G1)

table(GroupOS$Class, GroupOS$Group)
```

# immune distribution between 2 groups
# Function
```{r}
Plot_Cell_Type <- function(DF, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = Cell_Type, 
                                      y = Percentage, 
                                      by = Group, 
                                      color = Group)) +
                      geom_boxplot(position = position_dodge(width = 0.95),
                                   #outlier.shape = NA, 
                                   #outlier.colour = NA, 
                                   outlier.size = 0.25) +
                      geom_point(position = position_jitterdodge(
                                 dodge.width= 0.95,
                                 jitter.width = 0.25), 
                                 alpha=0.5,
                                 size = 0.1) +
                      stat_compare_means(aes(group = Group), 
                                        label = "p.format",
                                        label.x = P_X, 
                                        label.y = P_Y, 
                                        size = 3.6, 
                                        method = "wilcox.test") +
                      labs(x = "", y = "Cell Type Percent", 
                         colour = "") +
                      theme_classic() +
                      scale_color_nejm() +
                      theme(legend.position = "bottom",
                          text=element_text(size=12, face = "bold"),
                          plot.margin = unit(c(6,0,6,0), "pt")) + 
                  # legend.position controls where is legend
                      scale_y_continuous(labels = scales::percent_format(
                                       accuracy = 1)) 
                return(PLOT)
                
}

NLRPlot <- function(DF, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = Cell_Type, 
                                      y = Percentage, 
                                      by = Group, 
                                      color = Group)) +
                      geom_boxplot(position = position_dodge(width = 0.95),
                                   #outlier.shape = NA, 
                                   #outlier.colour = NA, 
                                   outlier.size = 0.25) +
                      geom_point(position = position_jitterdodge(
                                 dodge.width= 0.95,
                                 jitter.width = 0.25), 
                                 alpha=0.5,
                                 size = 0.1) +
                      stat_compare_means(aes(group = Group), 
                                        label = "p.format",
                                        label.x = P_X, 
                                        label.y = P_Y, 
                                        size = 3.6, 
                                        method = "wilcox.test") +
                      labs(x = "", y = "Ratio", 
                         colour = "") +
                      theme_classic() +
                      scale_color_nejm() +
                      theme(legend.position = "none")
                return(PLOT)
}
```

# Plot (Peripheral Blood)
```{r}
# cell proportion
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Bladder_Extended_cell_proportion.RData")
colnames(Extended2)[1] <- "Sample_ID"
GroupOS_2 <- GroupOS_2 %>% 
             select(-Bas,-Bmem,-Bnv,-CD4mem,-CD4nv,-CD8mem,-CD8nv,-Eos,-Mono,-Neu,-NK,-Treg) %>%
             left_join(Extended2) %>%
             mutate(T_cell = CD4mem + CD4nv + CD8mem + CD8nv + Treg, # not show this one
                     Lymph = Bmem + Bnv + NK + T_cell) %>%
              mutate(NLR = ifelse(Lymph == 0,Neu/0.01, Neu/Lymph))

Estimate1 <- GroupOS_2 %>%
             dplyr::select(Group, Bas, Bmem, Bnv, 
                           Eos, Mono, NK, Treg) %>%
             gather(key = Cell_Type, value = Percentage,-Group) %>%
             mutate(Percentage = Percentage/100)
Cell_Type_Plot = Estimate1 %>%
                 Plot_Cell_Type(DF = ., P_X = 0, P_Y = .22)

Estimate2 <- GroupOS_2 %>%
             dplyr::select(Group, CD4mem, CD4nv, CD8mem, CD8nv) %>%
             gather(key = Cell_Type, value = Percentage,-Group) %>%
             mutate(Percentage = Percentage/100)
Cell_Type_Plot2 = Estimate2 %>%
                  Plot_Cell_Type(DF = ., P_X = 0, P_Y = .4)

Estimate3 <- GroupOS_2 %>%
             dplyr::select(Group, Neu) %>%
             gather(key = Cell_Type, value = Percentage, -Group) %>%
             mutate(Percentage = Percentage/100)
Cell_Type_Plot3 = Estimate3 %>%
                  Plot_Cell_Type(DF = ., P_X = 0, P_Y = 1)

Estimate4 <- GroupOS_2 %>%
             dplyr::select( Group, NLR) %>%
             gather(key = Cell_Type, value = Percentage,-Group)
Cell_Type_Plot4 = Estimate4 %>%
                  NLRPlot(DF = ., P_X = 0, P_Y = 35)

# save
Row1 = plot_grid(Cell_Type_Plot, ncol = 1)
Row2 = plot_grid(Cell_Type_Plot3,Cell_Type_Plot2, Cell_Type_Plot4, ncol = 3, rel_widths = c(1,3,1))
Comb_Plot = plot_grid(Row1, Row2, ncol = 1, rel_heights = c(1.1,1))
ggsave(filename = paste0("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/Survival tree/NMIBC_Blood.pdf"), plot = Comb_Plot, height = 9, width = 7.5)
```