---
title: "SS_RPMM_OS_Hannum_acc"
author: "Ji-Qing Chen"
date: "2/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RPMM)
library(survival)
library(survminer)
library(nlme)
library(ggsci)
library(matrixStats)

library(dplyr)
library(tidyverse)
library(ggplot2)
source("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/SS_RPMM_function_limit_to2.R")

'%!in%' <- Negate('%in%')
```

# load data
```{r}
# beta
beta <- readRDS("/Users/chenjiqing/Public/bladder cancer/raw data/AllBC_Combat_Betas2.rds")
dim(beta)

# phenotype
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/Age_acc/Ctic_Project/AllBC_Age_acc2.RData") # immune cell already winsorized
```

# Exclude unnecessary CpGs
```{r}
annot <- readRDS("/Users/chenjiqing/Public/bladder cancer/GSEA and LOLA and GenomicContext/EPIC.hg19.manifest.rds")
annot<- as.data.frame(annot) %>% rownames_to_column(.)
table(annot$MASK_general)
row_contain_rs <- annot %>% filter(grepl("rs", .$rowname))
row_contain_ch <- annot %>% filter(grepl("ch", .$rowname))
annot_filtered <- annot %>% 
                  filter(rowname %!in% row_contain_rs$rowname) %>%
                  filter(rowname %!in% row_contain_ch$rowname) %>%
                  filter(MASK_general == "FALSE") %>%
                  filter(seqnames != "chrX") %>% 
                  filter(seqnames != "chrY")
rm(row_contain_rs,row_contain_ch)

beta2 <- as.data.frame(beta) %>% rownames_to_column(.)
rm(beta,annot)
beta.filtered <- beta2 %>% 
                 filter(rowname %in% annot_filtered$rowname)
rm(beta2,annot_filtered)
beta.filtered <- beta.filtered %>% column_to_rownames(., var = "rowname") %>% as.matrix(.)
dim(beta.filtered) # 726856 CpGs would be input
```

# Exclude unnecessary CpGs (New)
```{r}
test <- read.table(file = '/Users/chenjiqing/Public/bladder cancer/Tumor FFPE/First batch/Analysis/Cluster/EPIC.hg19.manifest.tsv', sep = '\t', header = TRUE)
row_contain_rs <- test %>% filter(grepl("rs", .$probeID))
row_contain_ch <- test %>% filter(grepl("ch", .$probeID))
annot_filtered <- test %>% 
                  filter(probeID %!in% row_contain_rs$probeID) %>%
                  filter(probeID %!in% row_contain_ch$probeID) %>%
                  filter(MASK_general == "FALSE") %>%
                  filter(CpG_chrm != "chrX") %>% 
                  filter(CpG_chrm != "chrY")
```

# clean data
```{r}
# 10% with the most variability in methylation beta value across all samples.
k <- nrow(beta.filtered)*0.1
sele <- order(rowVars(beta.filtered), decreasing=TRUE)[1:k]
beta2 <- beta.filtered[sele, ]
rm(beta.filtered)

NMIBC_immune_Age <- AllBC_immune_Age %>% filter(muscinv == "no")

a <- as.character(NMIBC_immune_Age$Sample_ID)
NMIBC_immune_Age <- NMIBC_immune_Age %>% column_to_rownames(., var = 'Sample_ID')
beta_final <- beta2[,a]
rm(beta2)
beta_final <- t(beta_final)
View(beta_final[1:5,1:5])

P1 = dim(NMIBC_immune_Age)[2]

```

# STEP 1: Randomly split the full data into training and testing sets.
```{r}
BLCR_Split = TrainTestSplit(beta_final, NMIBC_immune_Age, Strat = "TenDead", seed =1, propTrain = 2/3)

BLCR_TrainingData = BLCR_Split[[1]]
BLCR_TestingData = BLCR_Split[[2]]

# Extract the beta values and the covariate values from the training and testing sets from previous STEP.
BLCR_Training_Betas = BLCR_TrainingData[,-(1:P1)]
BLCR_Training_Covariates = BLCR_TrainingData[,(1:P1)]

BLCR_Testing_Betas = BLCR_TestingData[,-(1:P1)]
BLCR_Testing_Covariates = BLCR_TestingData[,(1:P1)]

rm(BLCR_Split, BLCR_TrainingData, BLCR_TestingData)
```

# STEP 2: Identify the CpG loci that are most associated with the clinical variable of interest using only the training data. If the clinical variable of interest is time to event data (i.e. survival or disease recurrence) then use the function called MostImpCpGsSurvival.
```{r}
BLCR_OS_Scores = MostImpCpGsSurvival(Y = BLCR_Training_Betas,
          covariates = BLCR_Training_Covariates, times = "TenYearSurv", censor = "TenDead",
          terms = c("Age","hannum_acc", "Sex", "Tumor_grade", "Smoking_status", "BCG"
                    ,"Bmem","CD4mem","CD8mem","CD8nv","Neu","Treg","NK","Bas"), 
          factors = c("Sex", "Tumor_grade", "Smoking_status", "BCG"))

```

# STEP 3a: Determine the number of high ranking CpG loci from the previous step to be used in fitting RPMM to the Training data.
```{r}
BLCRXvalidationResults4 = NestedXValidationSurvival(Y =  BLCR_Training_Betas, 
                                                   covariates = BLCR_Training_Covariates,
                                                   CoxScores = BLCR_OS_Scores, 
                                                   times = "TenYearSurv", 
                                                   censor = "TenDead", 
                                                   mrange =c(5,50),
                                                   method = "beta", 
                                                   L = 5, seeds = 1:5)
```

# STEP 3b:The selection of M can be based on what value of m yeilded the lowest median p-value as indicated in the HNSCCXvalidationResults and MesoXvalidationResults objects.
```{r}
# a better approach to select M would be to smooth the BLCRXvalidationResults object and chose the value of M where the smooth function of median p-values attains it's minimum value
mrange = 1:46
loessCurve = loess.smooth(mrange, BLCRXvalidationResults4, degree = 14)
MOpt_Bladder = subset(data.frame(loessCurve$x, loessCurve$y), loessCurve$y ==
min(loessCurve$y))[[1]]
par(mar = c(5,5,4,2))
plot(mrange, BLCRXvalidationResults4, cex = 0.75, xlab = "Number of top ranking loci (M)",
ylab = "Median P-value", cex.lab = 2, cex.axis = 1.5)
lines(loessCurve$x, loessCurve$y, lwd = 5)
abline(v = MOpt_Bladder, col = "red", lwd = 2, lty = "dashed")
```

# STEP 4: The final step of SS-RPMM is to fit an RPMM to the Training Data using the M CpG loci with the largest absolute T-Score, where M is determined based on the results from step 3.
```{r}
# Based on this solution, we want to predict the methylation class membership for the observations in the test data.  We achieve this using the PredMethClasses function.
BLCRMethClassesTesting = PredMethClasses(Ytrain = BLCR_Training_Betas,
                                         Ytest = BLCR_Testing_Betas, 
                                         Scores = BLCR_OS_Scores, 
                                         M = 41, 
                                         method = "beta")

BLCRMethClassesTraining = PredMethClasses(Ytrain = BLCR_Training_Betas,
                                         Ytest = BLCR_Training_Betas, 
                                         Scores = BLCR_OS_Scores, 
                                         M = 41, 
                                         method = "beta")

BLCR_classed1 <- BLCR_Testing_Covariates %>% mutate(Class = BLCRMethClassesTesting, SplitSet = "Test")
BLCR_classed2 <- BLCR_Training_Covariates %>% mutate(Class = BLCRMethClassesTraining, SplitSet = "Train")
BLCR_classed_cellprop <- rbind(BLCR_classed1,BLCR_classed2)
save(BLCR_OS_Scores, BLCRXvalidationResults4, BLCR_classed_cellprop, file = "/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/BLCR_classed_cellprop_5FoldCross.RData")
```

## POST SS-RPMM ANALYSES

# Following the prediction of the methylation classes in the Testing data, we now want to determine whether or not the methylation classes we've identified are clinically relevant.
```{r}
# We do this by testing the association between the predicted methylation classes in the test data and the clinical outcome of interest

# Kaplan-Meier survival plot by predicted methylation class in the testing data
plot(survfit(Surv(TenYearSurv,TenDead)~ BLCRMethClassesTesting, data = BLCR_Testing_Covariates),
      col = rainbow(length(levels(BLCRMethClassesTesting))),
      xlab = "Time (months)", ylab = "Probability of Survival", cex.lab = 1.5,
      cex.axis = 1.2, lwd = 2)
legend("bottomleft", legend = levels(BLCRMethClassesTesting),
       col = rainbow(length(levels(BLCRMethClassesTesting))),
       cex = 1.2, lty = 1, lwd = 2)
  
# Test whether or not the predicted methylation classes in the test data are associated with survival time (i.e. log-rank test).
survdiff(Surv(TenYearSurv,TenDead)~BLCRMethClassesTesting, data = BLCR_Testing_Covariates)

# Test whether or not the predicted methylation classes in the test data are associated with survival time, controlling for confounders.
CoxModBLCA = coxph(Surv(TenYearSurv,TenDead)~BLCRMethClassesTesting + Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Bnv + CD4mem + CD4nv + CD8mem + NK + Neu , data = BLCR_Testing_Covariates)
summary(CoxModBLCA)


# Load
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/BLCR_classed_cellprop_5FoldCross.RData")
# Re-plot KM plot
Test_data <- BLCR_classed_cellprop %>%
             filter(SplitSet == "Test")

fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Class, data = Test_data)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival, %",
  #ylim = c(0.3,1),
  legend = "top" ,legend.title = "", 
  legend.labs = c("rL","rR"),
  risk.table = TRUE, # For JAMA
  palette = "jama",
  #conf.int = T,
  #surv.median.line = "hv"
  risk.table.y.text.col=F, # For JAMA
  risk.table.title="No. at risk", # For JAMA
  risk.table.height = 0.18
  )

# Adjust y scale and table for JAMA
OS_10y$plot <- OS_10y$plot +
               scale_y_continuous(breaks = seq(0,1,by=0.2), labels = seq(0,100,by=20)) +
               theme(panel.grid.major.y = element_line(color = "grey",linetype = 1),
                     legend.position=c(0.08, 0.2),
                     legend.background = element_rect(linetype="solid", colour ="black"),
                     legend.title = element_blank(),
                     legend.text = element_text(size = 8),
                     axis.title.y = element_text(vjust = -3,margin = margin(r = 15)))
OS_10y$table <- OS_10y$table + 
                theme_cleantable() +
                theme(plot.title = element_text(hjust = -0.125,size=11))

surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Class, data = Test_data)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value
summary(os <- coxph(Surv(TenYearSurv, TenDead) ~ Class, data=Test_data))
OS_10y$plot <- OS_10y$plot + 
               annotate("text", x = 40, y = 0.3, label = " rR (referent group: rL), P = 1.96E-6; \n HR = 0.26 (95% CI, = 0.14-0.47)") +
               theme(plot.title = element_text(size = 5))
# size 5x8
```

# For whole data
```{r}
# Kaplan-Meier survival plot by predicted methylation class in the testing data
plot(survfit(Surv(TenYearSurv,TenDead)~ Class, data = BLCR_classed_cellprop),
      col = rainbow(length(levels(BLCR_classed_cellprop$Class))),
      xlab = "Time (months)", ylab = "Probability of Survival", cex.lab = 1.5,
      cex.axis = 1.2, lwd = 2)
legend("bottomright", legend = levels(BLCR_classed_cellprop$Class),
       col = rainbow(length(levels(BLCR_classed_cellprop$Class))),
       cex = 1.2, lty = 1, lwd = 2)
  
# Test whether or not the predicted methylation classes in the test data are associated with survival time (i.e. log-rank test).
survdiff(Surv(TenYearSurv,TenDead)~Class, data = BLCR_classed_cellprop)

# Test whether or not the predicted methylation classes in the test data are associated with survival time, controlling for confounders.
CoxModBLCA = coxph(Surv(TenYearSurv,TenDead)~Class + Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG + Bnv + CD4mem + CD4nv + CD8mem + NK + Neu , data = BLCR_classed_cellprop)
summary(CoxModBLCA)

# Load
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/BLCR_classed_cellprop_5FoldCross.RData")
# Re-plot KM plot
fit_10y = survfit(Surv(TenYearSurv, TenDead) ~ Class, data = BLCR_classed_cellprop)
OS_10y <- ggsurvplot(
  fit_10y,
  xlab = "Months", 
  ylab = "10-year overall survival, %",
  #ylim = c(0.3,1),
  legend = "top" ,legend.title = "", 
  legend.labs = c("rL","rR"),
  risk.table = TRUE, # For JAMA
  palette = "jama",
  #conf.int = T,
  #surv.median.line = "hv"
  risk.table.y.text.col=F, # For JAMA
  risk.table.title="No. at risk", # For JAMA
  risk.table.height = 0.18
  )

# Adjust y scale and table for JAMA
OS_10y$plot <- OS_10y$plot +
               scale_y_continuous(breaks = seq(0,1,by=0.2), labels = seq(0,100,by=20)) +
               theme(panel.grid.major.y = element_line(color = "grey",linetype = 1),
                     legend.position=c(0.08, 0.2),
                     legend.background = element_rect(linetype="solid", colour ="black"),
                     legend.title = element_blank(),
                     legend.text = element_text(size = 8),
                     axis.title.y = element_text(vjust = -3,margin = margin(r = 15)))
OS_10y$table <- OS_10y$table + 
                theme_cleantable() +
                theme(plot.title = element_text(hjust = -0.125,size=11))


surv_median(fit_10y)
surv_diff <- survdiff(Surv(TenYearSurv, TenDead) ~ Class, data = BLCR_classed_cellprop)
pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = FALSE) # log-rank test p value
summary(os <- coxph(Surv(TenYearSurv, TenDead) ~ Class, data=BLCR_classed_cellprop))
OS_10y$plot <- OS_10y$plot + 
               annotate("text", x = 40, y = 0.3, label = " rR (referent group: rL), P = 2.37E-13; \n HR = 0.31 (95% CI, = 0.23-0.44)") +
               theme(plot.title = element_text(size = 10))
```

# Cox not include immune
```{r}
CoxModBLCA = coxph(Surv(TenYearSurv,TenDead)~BLCRMethClassesTesting + Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG, data = BLCR_Testing_Covariates)
summary(CoxModBLCA)

CoxModBLCA = coxph(Surv(TenYearSurv,TenDead)~Class + Age + hannum_acc + Sex + Tumor_grade + Smoking_status + BCG, data = BLCR_classed_cellprop)
summary(CoxModBLCA)
```

______________________________________________________________
# Finding optimal CpGs
```{r}
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/BLCR_classed_cellprop_5FoldCross.RData")

MOpt_Bladder+5-1 # 41 CpGs

optimal_CpG <- rownames_to_column(BLCR_OS_Scores, var = "CpG")
optimal_CpG <- optimal_CpG[1:41,]

# add annotation
annEPIC <- getAnnotation("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annEPIC <- as.data.frame(annEPIC) %>% rownames_to_column(var = "CpG")
ann <- annEPIC %>% 
       select(CpG,Phantom5_Enhancers,DNase_Hypersensitivity_Evidence_Count,
              Relation_to_Island,chr,UCSC_RefGene_Name,UCSC_RefGene_Group) %>%
       mutate(DHS = ifelse(DNase_Hypersensitivity_Evidence_Count == "","No","Yes"),
              Enhancer = ifelse(Phantom5_Enhancers == "","No","Yes")) %>%
       select(CpG,DHS,Enhancer,Relation_to_Island,
              chr,UCSC_RefGene_Name,UCSC_RefGene_Group)

optimal_CpG <- optimal_CpG %>% left_join(., ann)

# Previous paper CpG
load("/Users/chenjiqing/Public/bladder cancer/EWAS/RFS/EWAS_strata_winsor.RData")

DMPs_0.005_cell <- DMPs_cell$table %>%
                   rownames_to_column(., var = "CpG") %>%
                   filter(p.value < 0.005)

DMPs_0.005_cell2 <- DMPs_0.005_cell %>%
                    select(CpG, estimate, p.value, p.adjust) %>%
                    left_join(.,ann)

# Find Common CpGs
common_CpG <- intersect(DMPs_0.005_cell2$CpG,optimal_CpG$CpG)
common_CpG <- optimal_CpG %>% filter(CpG %in% common_CpG)
```

# Heatmap for 41 methylation level
```{r}
library(pheatmap)
library(RColorBrewer)
#beta <- readRDS("/Users/chenjiqing/Public/bladder cancer/raw data/AllBC_Combat_Betas2.rds")
dim(beta)
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/BLCR_classed_cellprop_5FoldCross.RData")

# Target CpG
MOpt_Bladder+5-1 # 41 CpGs

optimal_CpG <- rownames_to_column(BLCR_OS_Scores, var = "CpG")
optimal_CpG <- optimal_CpG[1:41,]

a <- as.character(rownames(BLCR_classed_cellprop))
beta <- beta[,a]
b <- as.character(optimal_CpG$CpG)
beta_for41CpGs <- beta[b,]
e <- as.character(rownames(BLCR_classed_cellprop %>% filter(Class == "rR")))
f <- as.character(rownames(BLCR_classed_cellprop %>% filter(Class == "rL")))
beta_for41CpGs_rR <- beta_for41CpGs[,e]
beta_for41CpGs_rL <- beta_for41CpGs[,f]
beta_for41CpGs_Mean_rR <- beta_for41CpGs_rR %>%
                          as.data.frame() %>%
                          mutate(rR = rowMeans(beta_for41CpGs_rR)) %>%
                          rownames_to_column(., var = "CpG") %>%
                          select(CpG,rR)
beta_for41CpGs_Mean_rL <- beta_for41CpGs_rL %>%
                          as.data.frame() %>%
                          mutate(rL = rowMeans(beta_for41CpGs_rL)) %>%
                          rownames_to_column(., var = "CpG") %>%
                          select(CpG,rL)
beta_for41CpGs_Mean <- beta_for41CpGs_Mean_rL %>%
                       left_join(beta_for41CpGs_Mean_rR) %>%
                       column_to_rownames(., var = "CpG") %>%
                       as.matrix()

library(minfi)
annEPIC <- getAnnotation("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annEPIC <- as.data.frame(annEPIC) %>% rownames_to_column(var = "CpG")

# Find 41 CpGs
annEPIC_41 <- annEPIC %>% 
              filter(Name %in% b) %>%
              select(rowname,Phantom5_Enhancers,DNase_Hypersensitivity_Evidence_Count,Relation_to_Island) %>%
              mutate(DHS = ifelse(DNase_Hypersensitivity_Evidence_Count == "","No","Yes"),
              Enhancer = ifelse(Phantom5_Enhancers == "","No","Yes")) %>%
              select(rowname,DHS,Enhancer,Relation_to_Island)
colnames(annEPIC_41) <- c("CpG","DHS","Enhancer","Relation_to_Island")
annEPIC_41 <- column_to_rownames(annEPIC_41, var = "CpG")

# Annotation for 41 CpGs
ann <- annEPIC %>% 
       filter(Name %in% b) %>%
       select(CpG,Phantom5_Enhancers,DNase_Hypersensitivity_Evidence_Count,
              Relation_to_Island,chr,UCSC_RefGene_Name,UCSC_RefGene_Group) %>%
       mutate(DHS = ifelse(DNase_Hypersensitivity_Evidence_Count == "","No","Yes"),
              Enhancer = ifelse(Phantom5_Enhancers == "","No","Yes")) %>%
       select(CpG,DHS,Enhancer,Relation_to_Island,
              chr,UCSC_RefGene_Name,UCSC_RefGene_Group)
optimal_CpG_annotation <- optimal_CpG %>% left_join(., ann) %>% select(-AbsCoxScore)
#write_csv(optimal_CpG_annotation ,"/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/optimal_41CpG_annotation.csv")

# phenotype
phenotype <- BLCR_classed_cellprop %>%
             mutate(Age = ifelse(Age <= 65, "Age <= 65", "Age > 65"),
                    Hannum_avv = ifelse(hannum_acc <= median(hannum_acc), "Low Age_acc", "High Age_acc"),
                    NLR = ifelse(NLR <= median(NLR), "Low NLR", "High NLR"),
                    BCG = ifelse(BCG == "With Immuno","With BCG","Without BCG"),
                    Event = ifelse(TenDead == 1, "Deceased","Alive")) %>%
             select(Age,Hannum_avv,Sex,Tumor_grade,Smoking_status,BCG,Event,NLR,Class,SplitSet)

# For testing set
phenotype_test <- phenotype %>%
                  filter(SplitSet=="Test") %>%
                  select(-SplitSet)
c <- as.character(rownames(phenotype_test))
beta_for41CpGs_test <- beta_for41CpGs[,c]
```

# Save heatmap function
```{r}
save_pheatmap_png <- function(x, filename, width=1400, height=1600, res = 150) {
  png(filename, width = width, height = height, res = res)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}
```

# Heatmap
```{r}
mat_colors <- list(group = brewer.pal(2, "Set1"))

# For mean beta in train and test No row annotation
Top41_heat_mean <- pheatmap(
  beta_for41CpGs_Mean,
  show_rownames = F, 
  show_colnames = T,
  #cluster_rows = F,
  #annotation_col = phenotype_test,
  #annotation_row = annEPIC_37,
  annotation_colors = mat_colors,
  color = colorRampPalette(c("yellow", "blue"))(100), 
  border_color = NA,
  fontsize = 10, 
  fontsize_col = 10,
  main = "Mean Beta Values of Optimal 41 CpGs used in SS-RPMM [Whole NMIBC]"
)

save_pheatmap_png(Top41_heat_mean, "Top41_heat_mean_noRowAnnot")
```

______________________________________________________________

# immune cell type proportion difference
```{r}
# cell proportion
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Bladder_Extended_cell_proportion2.RData")
load("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/BLCR_classed_cellprop_5FoldCross.RData")

NMIBC_immune <- BLCR_classed_cellprop %>% 
                rownames_to_column(., var = 'Sample_Name') %>% 
                select(Sample_Name, Class) %>%
                left_join(Extended2) %>%
                mutate(T_cell = CD4mem + CD4nv + CD8mem + CD8nv + Treg, # not show this one
                       Lymph = Bmem + Bnv + NK + T_cell) %>%
                mutate(NLR = ifelse(Lymph == 0,Neu/0.01, Neu/Lymph)) %>%
                select(-T_cell,-Lymph)
```

# Function
```{r}
# Plot Cell Type
Plot_Cell_Type <- function(DF, YLab, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = Cell_Type, 
                                      y = Percentage,
                                      #color = Class,
                                      fill = Class,
                                      by = Class)) +
                       stat_boxplot(geom = "errorbar", width = 0.2,
                                    position = position_dodge(width = 0.95)) + # For JAMA
                       geom_boxplot(position = position_dodge(width = 0.95), 
                                    outlier.shape = 1, 
                                    #outlier.colour = NA, 
                                    outlier.size = 1) +
                      #geom_point(position = position_jitterdodge(dodge.width= 0.95, 
                       #                                          jitter.width = 0.25), 
                        #         alpha=0.5,size = 0.1) +
                      stat_compare_means(aes(group = Class), 
                                         label = "p.format", 
                                         label.x = P_X, label.y = P_Y, 
                                         size = 3.4, 
                                         method = "wilcox.test") +
                      labs(x = "", y = YLab, colour = "") +
                      theme_classic() +
                      #scale_color_nejm() +
                      scale_fill_jama() +
                      theme(legend.position = "bottom", # legend.position controls where is legend
                            panel.grid.major.y = element_line(color = "grey",linetype = 1), # For JAMA
                            legend.title = element_blank(), # For JAMA
                            text=element_text(size=12, 
                                              face = "bold"),
                            plot.margin = unit(c(6,0,6,0), "pt")) + 
                     scale_y_continuous(labels = scales::percent_format(accuracy = 1)) 
                return(PLOT)
}

NeuPlot <- function(DF){
            ggplot(DF, aes(x = "Neu",
                           y = Neu,
                           #color = Class,
                           fill = Class,
                           by = Class)) +
            stat_boxplot(geom = "errorbar", width = 0.2,
                         position = position_dodge(width = 0.95)) + # For JAMA
            geom_boxplot(position = position_dodge(width = 0.95), 
                         outlier.shape = 1, 
                         #outlier.colour = NA, 
                         outlier.size = 1.5) + 
            #geom_point(position = position_jitterdodge(dodge.width= 0.95, 
             #                                          jitter.width = 0.25),
              #         alpha=0.5,size = 0.1) + 
            stat_compare_means(aes(group = Class), 
                               label = "p.format", 
                               label.x = 0, label.y = 1, 
                               size = 3.4, 
                               method = "wilcox.test") +
            labs(x = "", y = "Cell Type Percent") +
            theme_classic() +
            #scale_color_nejm() +
            scale_fill_jama() +
            theme(legend.position = "none", # legend.position controls where is legend
                  panel.grid.major.y = element_line(color = "grey",linetype = 1), # For JAMA
                  text=element_text(size=12,face = "bold"),
                  plot.margin = unit(c(6,0,6,0), "pt")) +
            scale_y_continuous(labels = scales::percent_format(accuracy = 1))
}

# Plot Cell Type
Cell_Ratio_Plot <- function(DF, YLab, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = "NLR", 
                                      y = NLR, 
                                      #color = Class,
                                      fill = Class,
                                      by = Class)) +
                       geom_boxplot(position = position_dodge(width = 0.95), 
                                    outlier.shape = 1, 
                                    #outlier.colour = NA, 
                                    outlier.size = 1.5) +
                       #geom_point(position = position_jitterdodge(dodge.width= 0.95, 
                        #                               jitter.width = 0.25),
                         #        alpha=0.5,size = 0.1) +
                       stat_compare_means(aes(group = Class), 
                                          label = "p.format", 
                                          label.x = P_X, 
                                          label.y = P_Y, 
                                          size = 3.4, 
                                          method = "wilcox.test") +
                      labs(x = "", y = YLab, colour = "") +
                      theme_classic() +
                      #scale_color_nejm() +
                      scale_fill_jama() +
                      theme(legend.position = "none",
                            panel.grid.major.y = element_line(color = "grey",linetype = 1), # For JAMA
                            text=element_text(size=12, face = "bold"),
                            plot.margin = unit(c(6,0,6,0), "pt"), 
                            axis.text.x = element_text(size = 7))
                return(PLOT)
                
        }
```

# plot
```{r}
Estimate <- NMIBC_immune %>%
            dplyr::select(Class, Bas, Bmem, Bnv, Eos, Mono, NK, Treg) %>%
            gather(key = Cell_Type, value = Percentage, -Class) %>%
            mutate(Percentage = Percentage/100)
plot1 = Plot_Cell_Type(Estimate, YLab = "Cell Type Percent", P_X = 0, P_Y = .21)

Estimate2 <- NMIBC_immune %>%
             dplyr::select(Class, CD4mem, CD4nv, CD8mem, CD8nv) %>%
             gather(key = Cell_Type, value = Percentage, -Class) %>%
             mutate(Percentage = Percentage/100)
plot2 = Plot_Cell_Type(Estimate2, YLab = "Cell Type Percent", P_X = 0, P_Y = .4)

PlotNeu <- NMIBC_immune %>% mutate(Neu = Neu/100) %>% NeuPlot()

Ratios <- NMIBC_immune %>%
          dplyr::select(Class, NLR)
PlotNLR <- Cell_Ratio_Plot(Ratios,YLab = 'Ratio', P_X = 1.5, P_Y = 25)

# save
Row1 = plot_grid(plot1, ncol = 1)
Row2 = plot_grid(PlotNeu, plot2, PlotNLR, ncol = 3, rel_widths = c(1,3,1))
Comb_Plot = plot_grid(Row1, Row2, ncol = 1, rel_heights = c(1, 1))
ggsave(filename = paste0("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/plot/NMIBC_summary proportion_hannum_acc_JAMA.pdf"), plot = Comb_Plot, height = 9, width = 7.7)
```

______________________________________________________________

# age and age acceleration
```{r}
NMIBC_age <- BLCR_classed_cellprop %>% 
             select(Age, mAge_Hannum, hannum_acc, Class) %>%
             `colnames<-`(c('Age','Hannum_mAge', 'Age_Acceleration', 'Class'))
```

# Function
```{r}
# Plot Cell Type
Plot_Age_Type <- function(DF, YLab, P_X, P_Y) {
                
                PLOT = ggplot(DF, aes(x = Ages, 
                                      y = Values,
                                      #color = Class,
                                      fill = Class,
                                      by = Class)) +
                       stat_boxplot(geom = "errorbar", width = 0.2,
                                    position = position_dodge(width = 0.95)) + # For JAMA
                       geom_boxplot(position = position_dodge(width = 0.95), 
                                    outlier.shape = 1, 
                                    #outlier.colour = NA, 
                                    outlier.size = 1.3) +
                      #geom_point(position = position_jitterdodge(dodge.width= 0.95, 
                       #                                          jitter.width = 0.25), 
                        #         alpha=0.5,size = 0.1) +
                      stat_compare_means(aes(group = Class), 
                                         label = "p.format", 
                                         label.x = P_X, label.y = P_Y, 
                                         size = 3.4, 
                                         method = "wilcox.test") +
                      labs(x = "", y = YLab, colour = "") +
                      theme_classic() +
                      #scale_color_nejm() +
                      scale_fill_jama() +
                      theme(legend.position = "bottom", # legend.position controls where is legend
                            panel.grid.major.y = element_line(color = "grey",linetype = 1), # For JAMA
                            legend.title = element_blank(), # For JAMA
                            text=element_text(size=12, 
                                              face = "bold"),
                            plot.margin = unit(c(6,0,6,0), "pt"))
                return(PLOT)
}

```

# plot
```{r}
Estimate <- NMIBC_age %>%
            dplyr::select(Class, Age, Hannum_mAge) %>%
            gather(key = Ages, value = Values, -Class) 
plot1 = Plot_Age_Type(Estimate, YLab = "Years", P_X = 0, P_Y = 97)

Estimate2 <- NMIBC_age %>%
            dplyr::select(Class, Age_Acceleration) %>%
            gather(key = Ages, value = Values, -Class) 
plot2 = Plot_Age_Type(Estimate2, YLab = "Years", P_X = 0, P_Y = 24)

# save
Comb_Plot = plot_grid(plot1, plot2, ncol = 1, rel_heights = c(1, 1))
ggsave(filename = paste0("/Users/chenjiqing/Public/bladder cancer/Extended Deconvolution/Script/SS_RPMM/plot/NMIBC_age_and_mage_hannum_acc_JAMA.pdf"), plot = Comb_Plot, height = 9, width = 7.5)
```